From noise to knowledge: how randomness generates novel phenomena and information

### 2017-10-24 paper.Rmd
WIP

```{r include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
library(hrbrthemes)
library(ggthemes)
library(tidyverse)
#ggplot2::theme_set(hrbrthemes::theme_ipsum())
```

@@

**Chaos**: Deterministic chaos [@Desharnais2001] is another mechanism of introducing significant variation into a simple
model. Chaos can a look a lot like noise: indeed, it can be observationally indistinguishable from a stochastic process [@Werndl2009].  Despite substantial interest in previous decades and the relative ease of encountering chaos in simple ecological models (particularly in differences equations), and convincing reconstructions of chaotic dynamics in laboratory manipulations [@Constantino1997] the prevelance of chaotic dynamics as a driving factor in most ecological systems remains unclear. Meanwhile, the undeniable influences of

@@

```{r}
read_csv("appendix/quasicycles.csv") %>%
  gather(species, pop, -t, -sigma) %>%
  ggplot(aes(t, pop, col=species)) + geom_line() + 
  facet_wrap(~sigma) + scale_color_ptol() + 
  labs(x="Time", 
       y="Population density",
       title="Stochasticicity induces quasi-cycles",
       caption="Left panel: neglible  noise results only in damped oscillations.\nRight panel: small noise creates sustained oscillations.")
```

@@

```{r}
read_csv("appendix/noisy_switch.csv")  %>% ggplot(aes(t,x)) + 
  geom_line() + 
  labs(title="Stochastic switching", 
       subtitle = "Noise drives transitions between alternative stable states")
```

@@

p1 <- tipping %>%
  select(-x) %>%
  gather(series, value, -t, -tip_time) %>%
  ggplot(aes(t, value)) + geom_line() + 
  geom_vline(aes(xintercept = tip_time), col="red", lty=2) +
  facet_wrap(~series, ncol = 2, scales="free_y") +
  #theme_ipsum() +
  labs(title="A: Early warning signs")

p2 <- tipping %>% 
  select(t, population = x, tip_time) %>%
  ggplot(aes(t, population)) + geom_line() + 
  geom_vline(aes(xintercept = tip_time), col="red", lty=2) + 
  scale_y_log10() + 
  #theme_ipsum() + 
  labs(title="B: Population density",
       caption = "Vertical red dashed line indicates tipping point location")

library(cowplot)
ggdraw() +
  draw_plot(p2, x = 0, y = 0, width = 1, height = 0.5) +
  draw_plot(p1, x = 0, y = .4, width = 1, height = 0.5)
```

@@

>>>>

```{r include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
library(ggthemes)
library(tidyverse)
library(cowplot)
#library(hrbrthemes)
#extrafont::loadfonts()
#ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())
#hrbrthemes::import_roboto_condensed()

ggplot2::theme_set(theme_cowplot())
```

@@

**Chaos**: Deterministic chaos [@Desharnais2001] is another mechanism of introducing significant variation into a simple
model. Chaos can a look a lot like noise: indeed, it can be observationally indistinguishable from a stochastic process [@Werndl2009].  Despite substantial interest in previous decades and the relative ease of encountering chaos in simple ecological models (particularly in differences equations), and convincing reconstructions of chaotic dynamics in laboratory manipulations [@Costantino1997] the prevelance of chaotic dynamics as a driving factor in most ecological systems remains unclear. Meanwhile, the undeniable influences of

@@

```{r}
read_csv("appendix/quasicycles.csv") %>%
  gather(species, population, -t, -sigma) %>%
  rename(time = t) %>%
  ggplot(aes(time, population, col = species)) + geom_line() + 
  facet_wrap(~ sigma) + scale_color_ptol() #+ 
#  labs(x="Time", 
#       y="Population density",
#       title="Noise induces quasi-cycles",
#       caption="Left panel: neglible  noise results only in damped oscillations.\nRight panel: small noise creates sustained oscillations.")
#ggsave("figures/quasicycles.pdf")
```

@@

```{r}
read_csv("appendix/noisy_switch.csv")  %>% ggplot(aes(t,x)) + 
  geom_line() #+ 
  #labs(title="Stochastic switching", 
  #     subtitle = "Noise drives transitions between alternative stable states")
#ggsave("figures/noisy_switch.pdf")

```

@@

p1 <- tipping %>%
  select(-x) %>%
  gather(series, value, -t, -tip_time) %>%
  ggplot(aes(t, value)) + geom_line() + 
  geom_vline(aes(xintercept = tip_time), col="red", lty=2) +
  facet_wrap(~series, ncol = 2, scales="free_y") # +
#  theme_ipsum_rc() +
#  labs(title="A: Early warning signs")

p2 <- tipping %>% 
  select(t, population = x, tip_time) %>%
  ggplot(aes(t, population)) + geom_line() + 
  geom_vline(aes(xintercept = tip_time), col="red", lty=2) #+ 
#  scale_y_log10() # + 
#  theme_ipsum_rc() + 
#  labs(title="B: Population density",
#       caption = "Vertical red dashed line indicates tipping point location")

library(cowplot)
ggdraw() +
  draw_plot(p2, x = 0, y = 0, width = 1, height = 0.5) +
  draw_plot(p1, x = 0, y = .5, width = 1, height = 0.5)
ggsave("figures/tipping.pdf")
```

@@

## References
